Local Spectral Clustering of Density Upper Level Sets

### 2019-06-24 neurips/local_spectral_density_clustering_arxiv.tex
more updates -- updates to proof of Thm 1

\paragraph{Proof of Lemma \ref{lem: graph_conductance_profile_lb}}

@@

\section{Proof of Theorem \ref{thm: misclassification_rate}}
\label{sec: proof_of_misclassification_rate}
In this section, we prove both Theorem \ref{thm: misclassification_rate} and the relevant portions of Corollary \ref{cor: appr}.

Lemma \ref{lem: zhu} is a restatement of Lemma 3.4 of \cite{zhu2013}, translated into notation consistent with the rest of this paper. It is stated with respect to an arbitrary $S \subseteq G = (V,E)$, and a good set $S^g \subseteq S$ with $\vol(S^g; G) \geq \vol(S; G)/2$. \footnote{The sweep cuts of Lemma 3.4 of \cite{zhu2013} are stated directly with respect to $\vol(S;G)$, rather than $\vol_0$; this discrepancy leads to the extra factors of $1 - c$ and $1 + c$ which appear in our bounds as opposed to theirs.}

\begin{lemma}
	\label{lem: zhu}
	Let $\varepsilon \geq 0$, $v \in S^g$, $\alpha \leq \frac{1}{9\tau_{\infty}(G[S])}$, and $\vol_0 \in [1 - c,1 + c] \vol(S;G)$ for $0 < c < 1$. Consider taking sweep cuts of the approximate \pprspace vector $\pbf^{(\varepsilon)} := \pbf^{(\varepsilon)}(v,\alpha; G)$ of the form
	\begin{equation*}
	S_{\beta} := \left\{u \in V: \pbf^{(\varepsilon)} \geq \beta \frac{\deg(u; G)}{\vol_0} \right\}
	\end{equation*}
	For any $\beta < \frac{3}{5(1 + c)}$, and $0 \leq \varepsilon \leq \frac{1}{10} \vol(S;G)$ the following bounds hold:
	\begin{align}
	\vol(S_{\beta} \setminus S; G) & \leq \frac{2\Phi(A;G)}{\alpha \beta(1 - c)} \vol(S; G), \label{eqn: zhu_1}\\
	\vol(S \setminus S_{\beta}; G) & \leq \left(\frac{2\Phi(S;G)}{\alpha (\frac{3}{5} - \beta(1 + c))} + 8\Phi(S;G)\right)\vol(S; G).  \label{eqn: zhu_2}
	\end{align}
\end{lemma}

\paragraph{Proof of Theorem \ref{thm: misclassification_rate}.}

Observe that by Theorem \ref{thm: mixing_time_upper_bound} and the upper bounds $\alpha \leq \frac{1}{9 \tau_u(\theta)}, r < \sigma/2d$ given in \eqref{eqn: initialization}, with probability one as $n \to \infty$, $\alpha \leq \frac{1}{9\tau_{\infty}(G[S])}$. Moreover, 
\begin{enumerate}[\text{(i)}]
	\item we restrict $v \in \Csig[\Xbf]^g$
	\item the output of Algorithm \ref{alg: ppr} $\widehat{C} = S_{\beta}$ for some $1/40 < \beta < 1/11$
	\item by Lemma \ref{lem: ball_bounds_in_probability}, $\vol_0 \in [1/2,3/2]\vol_{n,r}(\Csig[\Xbf])$
	\item the \pprspace vector $\pbf(v,\alpha;G_{n,r})$ is simply an approximate \pprspace vector with $\varepsilon = 0$.
\end{enumerate}

All conditions of Lemma \ref{lem: zhu} are therefore satisfied, and consequently by \eqref{eqn: zhu_1}, \eqref{eqn: zhu_2}, Theorem \ref{thm: conductance_upper_bound}, and the lower bound $\alpha \geq 1/10\tau_u(\theta)$, the following bounds hold with probability tending to 1 as $n \to \infty$:
\begin{equation*}
\vol_{n,r}(\widehat{C} \setminus \Csig[\Xbf]) \leq 1600 \kappa(\Cset) \vol_{n,r}(\Csig[\Xbf]), \quad \textrm{and} \quad  \vol_{n,r}(\Csig[\Xbf] \setminus \widehat{C}) \leq 52 \kappa(\Cset) \vol_{n,r}(\Csig[\Xbf])
\end{equation*}
Noting that
\begin{align*}
\abs{\widehat{C} \setminus \Csig[\Xbf]} & \leq \frac{\vol_{n,r}(\widehat{C} \setminus \Csig[\Xbf])}{\degmin}   \\
\abs{\Csig[\Xbf] \setminus \widehat{C}} & \leq \frac{\vol_{n,r}(\Csig[\Xbf] \setminus \widehat{C})}{\degminwt}, \quad \textrm{and} \\
\abs{\Csig[\Xbf]} & \geq \frac{\vol_{n,r}(\Csig[\Xbf])}{\degmax}
\end{align*}
the claim follows in light of \eqref{eqn: min_degree_bound}, \eqref{eqn: min_cluster_degree_bound} and \eqref{eqn: max_degree_bound} from Lemma \ref{lem: ball_bounds_in_probability}.

@@

Let $s_t := q(\Wbf \Ibf{S})^t(\Wbf(\Ibf{S^c}))$ be the soakage vector out of $S \subseteq V$, where $\Ibf{S}$ is a $\abs{V} \times \abs{V}$ diagonal matrix with $(\Ibf{S})_{uu} = 1$ if $u \in S$ and 0 otherwise, and $\Ibf{S^c} := \Ibf{} - \Ibf{S}$. 
\begin{lemma}
	\label{lem: sum_of_soakages}
	Let $G = (V,E)$ be an unweighted, undirected graph with associated random walk matrix $\Wbf$. For any $T \in \mathbb{N}, T \geq 1$, $q \in \Reals^{\abs{V}}$, and $S \subseteq V$
	\begin{equation}
	\label{eqn: sum_of_soakages}
	q\Wbf^T = \sum_{t = 0}^{T - 1} s_t \Wbf^{T - t - 1} + q(\Wbf \Ibf{S})^T
	\end{equation}
	In particular, if $u \in V\setminus S$, then
	\begin{equation}
	\label{eqn: sum_of_soakages_2}
	q\Wbf^T(u) = \sum_{t = 0}^{T - 1} \left(s_t \Wbf^{T - t - 1}\right)(u)
	\end{equation}
\end{lemma}
\begin{proof}	
	We show \eqref{eqn: sum_of_soakages}, from which \eqref{eqn: sum_of_soakages_2} is an immediate consequence.

	To show \eqref{eqn: sum_of_soakages}, we proceed by induction on $T$. When $T = 1$,
	\begin{equation*}
	q\Wbf = q\Wbf\Ibf{S} + q\Wbf\Ibf{S^c} =  q\Wbf\Ibf{S} + s_0. 
	\end{equation*}

>>>>

\paragraph{Proof of Lemma \ref{lem: graph_conductance_profile_lb}:}

@@

\section{Proof of Theorem \ref{thm: misclassification_rate}}
\label{sec: proof_of_misclassification_rate}
In this section, we prove both Theorem \ref{thm: misclassification_rate}.
The result is in essence a consequence of the bounds in Theorem~\ref{thm: conductance_upper_bound} 
and Theorem~\ref{thm: mixing_time_upper_bound} bounding the normalized cut and mixing time, together with the result of \citet{zhu2013} relating the quality of the PPR output to these quantities, and the result in Lemma~\ref{lem: ball_bounds_in_probability}
that allows us to translate the guarantees on the volume to guarantees on the symmetric set difference.

% and the relevant portions of Corollary \ref{cor: appr}.
We begin with Lemma \ref{lem: zhu}, which is a restatement of Lemma 3.4 of \citet{zhu2013}, translated into notation consistent with the rest of this paper. It is stated with respect to an arbitrary $S \subseteq G = (V,E)$, and a good set $S^g \subseteq S$ with $\vol(S^g; G) \geq \vol(S; G)/2$.\footnote{The sweep cuts of Lemma 3.4 of \cite{zhu2013} are stated directly with respect to $\vol(S;G)$, rather than $\vol_0$; this discrepancy leads to the extra factors of $1 - c$ and $1 + c$ which appear in our bounds as opposed to theirs.}

\begin{lemma}
	\label{lem: zhu}
	Let $\varepsilon \geq 0$, $v \in S^g$, $\alpha \leq \frac{1}{9\tau_{\infty}(G[S])}$, and $\vol_0 \in [1 - c,1 + c] \vol(S;G)$ for $0 < c < 1$. Consider taking sweep cuts of the approximate \pprspace vector $\pbf^{(\varepsilon)} := \pbf^{(\varepsilon)}(v,\alpha; G)$ of the form
	\begin{equation*}
	S_{\beta} := \left\{u \in V: \pbf^{(\varepsilon)} \geq \beta \frac{\deg(u; G)}{\vol_0} \right\}
	\end{equation*}
	For any $\beta < \frac{3}{5(1 + c)}$, and $0 \leq \varepsilon \leq \frac{1}{10} \vol(S;G)$ the following bounds hold:
	\begin{align}
	\vol(S_{\beta} \setminus S; G) & \leq \frac{2\Phi(A;G)}{\alpha \beta(1 - c)} \vol(S; G), \label{eqn: zhu_1}\\
	\vol(S \setminus S_{\beta}; G) & \leq \left(\frac{2\Phi(S;G)}{\alpha (\frac{3}{5} - \beta(1 + c))} + 8\Phi(S;G)\right)\vol(S; G).  \label{eqn: zhu_2}
	\end{align}
\end{lemma}
%\paragraph{Proof of Theorem \ref{thm: misclassification_rate}.}
With this Lemma in place we
observe that by Theorem \ref{thm: mixing_time_upper_bound} and the upper bounds $\alpha \leq \frac{1}{9 \tau_u(\theta)}, r < \sigma/2d$ given in \eqref{eqn: initialization}, with probability one as $n \to \infty$, $\alpha \leq \frac{1}{9\tau_{\infty}(G[S])}$. Moreover, 
\begin{enumerate}
	\item we restrict $v \in \Csig[\Xbf]^g$
	\item the output of Algorithm \ref{alg: ppr} $\widehat{C} = S_{\beta}$, for some $1/40 < \beta < 1/11$
	\item by Lemma \ref{lem: ball_bounds_in_probability}, $\vol_0 \in [1/2,3/2]\vol_{n,r}(\Csig[\Xbf])$
	\item the \pprspace vector $\pbf(v,\alpha;G_{n,r})$ is simply an approximate \pprspace vector with $\varepsilon = 0$.
\end{enumerate}

All conditions of Lemma \ref{lem: zhu} are therefore satisfied, and consequently by \eqref{eqn: zhu_1}, \eqref{eqn: zhu_2}, Theorem \ref{thm: conductance_upper_bound}, and the lower bound $\alpha \geq 1/10\tau_u(\theta)$, the following bounds hold with probability tending to 1 as $n \to \infty$:
\begin{equation*}
\vol_{n,r}(\widehat{C} \setminus \Csig[\Xbf]) \leq 1600 \kappa(\Cset) \vol_{n,r}(\Csig[\Xbf]), \quad \textrm{and} \quad  \vol_{n,r}(\Csig[\Xbf] \setminus \widehat{C}) \leq 52 \kappa(\Cset) \vol_{n,r}(\Csig[\Xbf])
\end{equation*}
Noting that
\begin{align*}
\abs{\widehat{C} \setminus \Csig[\Xbf]} & \leq \frac{\vol_{n,r}(\widehat{C} \setminus \Csig[\Xbf])}{\degmin}   \\
\abs{\Csig[\Xbf] \setminus \widehat{C}} & \leq \frac{\vol_{n,r}(\Csig[\Xbf] \setminus \widehat{C})}{\degminwt}, \quad \textrm{and} \\
\abs{\Csig[\Xbf]} & \geq \frac{\vol_{n,r}(\Csig[\Xbf])}{\degmax}
\end{align*}
the theorem follows in light of \eqref{eqn: min_degree_bound}, \eqref{eqn: min_cluster_degree_bound} and \eqref{eqn: max_degree_bound} from Lemma \ref{lem: ball_bounds_in_probability}.

@@

Let $s_t := q(\Wbf \Ibf{S})^t(\Wbf(\Ibf{S^c}))$ be the soakage vector out of $S \subseteq V$, where $\Ibf{S}$ is a $\abs{V} \times \abs{V}$ diagonal matrix with $(\Ibf{S})_{uu} = 1$ if $u \in S$ and 0 otherwise, and $\Ibf{S^c} := \Ibf{} - \Ibf{S}$. 
\begin{lemma}
	\label{lem: sum_of_soakages}
	Let $G = (V,E)$ be an unweighted, undirected graph with associated random walk matrix $\Wbf$. For any $T \in \mathbb{N}, T \geq 1$, $q \in \Reals^{\abs{V}}$, and $S \subseteq V$
	\begin{equation}
	\label{eqn: sum_of_soakages}
	q\Wbf^T = \sum_{t = 0}^{T - 1} s_t \Wbf^{T - t - 1} + q(\Wbf \Ibf{S})^T
	\end{equation}
	In particular, if $u \in V\setminus S$, then
	\begin{equation}
	\label{eqn: sum_of_soakages_2}
	q\Wbf^T(u) = \sum_{t = 0}^{T - 1} \left(s_t \Wbf^{T - t - 1}\right)(u)
	\end{equation}
\end{lemma}
\begin{proof}	
	We show \eqref{eqn: sum_of_soakages}, from which \eqref{eqn: sum_of_soakages_2} is an immediate consequence.
	To show \eqref{eqn: sum_of_soakages}, we proceed by induction on $T$. When $T = 1$,
	\begin{equation*}
	q\Wbf = q\Wbf\Ibf{S} + q\Wbf\Ibf{S^c} =  q\Wbf\Ibf{S} + s_0. 
	\end{equation*}
